Recalibrating Gravitational Wave Phenomenological Waveform Model

### 2023-06-22 src/tex/ms.tex
Start working on discussion
added citation and fixed some comment

Many data analysis tasks in today's gravitational wave (GW) astrophysics, such
as match filtering search \citep{fill} and parameter estimation\citep{fill}, rely on accurate waveform
models. Because using numerical relativity (NR) waveforms in these tasks is
prohibitively expensive, the community has constructed waveform approximants
that can be evaluated much faster. There are three families of commonly used GW
approximants, which are effective-one-body (EOB) \citep{ossokine2020multipolar,
cotesta2020frequency, taracchini2014effective}, Numerical Relativity (NR)
surrogate \citep{islam2022surrogate, varma2019surrogate, varma2019surrogate2}
and phenomenological (Phenom) models \citep{husa2016frequency,
khan2016frequency, garcia2020multimode, pratten2021computationally}. While the
detail of construction of these models are different, they all have a set of
internal parameters that can be calibrated to NR waveforms. The quality of a
waveform model is determined by the ansatz used and the accuracy of the
calibrated parameters.

@@

The LIGO-VIRGO-KAGRA (LVK) collaboration
(\citep{LIGOScientific:2014pky,LIGOScientific:2021usb,LIGOScientific:2021djp,
VIRGO:2014yos,KAGRA:2020tym}.) has recently started their fourth observational
run on May 26, 2023. Because of the sensitivity improvement, the new run is
expected to double the size of current binary black hole (BBH) observations
\citep{abbott2020prospects}. The improved sensitivity also implies we expect to
detect individual events with a higher signal-to-noise ratio (SNR). This means
we can resolve more features in the signal, at the same time put a more
stringent requirement on the accuracy of our waveform model\citep{fill}.

@@

Because of the large number of calibration parameters (often in a couple hundreds if not
more), waveform models is usually calibrated in pieces \citep{fill}. This
ignores the correlation between different parts of the waveform model and limits
it quality. Recently, there has been efforts in rebuilding waveform models
\citep{fill} that supports automatic differentiation (AD)\citep{fill}, which is
a technique used to compute derivatives of functions down to machine precision
without issues of scaling up to high dimension or expression swelling. In
particular, {\ripple} \citep{ripple} exposes the calibration parameters to the
user. This allows us to make use of infrastructures that are heavily used in
machine learning, such as gradient descent and back propagation\cite{fill}, to
improve the calibration of the waveform models.

% Put more focus on the joint optimization instead of having a set of more accurate 
% coefficients, since the method is the main point
In this paper, we investigate the possibility of further improving the accuracy
of a waveform model, IMRPhenomD\citep{fill}, by jointly optimizing all the
calibration coefficients given a set of NR waveforms. Training on a few NR
waveforms, we demonstrate one can improve the match between IMRPhenomD and NR
waveforms over a decently sized parameter space, up to mass ratio $q=8$ We also
explore how different parts of the source parameter space (e.g. the primary and
second spins) respond to the optimization procedure, by optimizing the waveform
separately for different regions of the parameter space. This gives us hints on
whether the waveform model performs equally well in different regions of the
parameter space.

@@

In practice the $\Lambda$ parameters are fit for each section independently i.e., intermediate coefficients are fit whilst ignoring the MR region.
Finally, to map the grid of tuned $\Lambda$ parameters back to the intrinsic parameter space, IMRhenomD uses the polynomial function:
\begin{align} \label{eq:Lambda}
	\Lambda^i&=\lambda_{00}^i+\lambda_{10}^i\eta \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)(\lambda_{01}^i+\lambda_{11}^i\eta+\lambda_{21}^i\eta^2) \nonumber \\ 
	&+(\chi_{\mathrm{PN}}-1)^2(\lambda_{02}^i+\lambda_{12}^i\eta+\lambda_{22}^i\eta^2) \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)^3(\lambda_{03}^i+\lambda_{13}^i\eta+\lambda_{23}^i\eta^2)\,,
\end{align}
where the $\lambda$'s are the fitting coefficients we are going to optimize below, $\eta$ is
the symmetric mass ratio, and $\chi_{\mathrm{PN}}$ is the post-Newtonian spin
parameter, which is defined as 
\begin{align}
	\chi_{\mathrm{PN}}=\frac{m_1\chi_1+m_2\chi_2}{m_1+m_2}-\frac{38\eta}{113}(\chi_1+\chi_2)\,.
\end{align}
Here, $m_{1,2}$ and $\chi_{1,2}$ are the primary and secondary mass and spin,
respectively.

@@

%\kw{Need to work on this paragraph}
Although initially independent, the stiching procedure means that each section
of the waveform intrinsically depends on the full set of $\lambda$'s. 
A slightly inaccurate set of $\lambda$'s can therefore lead to inaccuracies in
the generated waveforms. 
Thus, the calibration of these coefficients is crucial to the accuracy
of IMRPhenom GW models. 
Importantly, since the fitting was performed on the individual segments,
the final waveform is not guarenteed to have $\lambda$'s close to global minima.

@@

At the time of construction this piece-wise approach was necessary since
$\lambda$ has 209 components, making the fitting to NR simulations computationally prohibitive.
Here, for the first time we recalibrate the $\lambda$ coefficients jointly. 
This is made possible by the use of gradient-based optimization algorithms,
enabled by AD from \jax\, and {\ripple}, which are significatnly more efficient in high dimensions.
% , gradients of IMRPhenomD can be easily obtained, thus
% allowing the use of gradient-based algorithms for us to recalibrate the model.

@@

Since we wish to optimize the model over the whole parameter space, we need to compare multiple model generated waveforms with NR waveforms. However, mismatch only quantifies the difference between IMRPhenomD and NR waveform for one particular set of intrinsic parameters. To take into account of various different waveforms in the parameter space, we pick waveforms from the different parts of the parameter space. We define the loss function as an average of training waveforms in two ways, the simple average of mismatches and the normalize average of mismatches,  
\begin{align}
	\mathcal{L}_{\mathrm{mean}}&=\frac{1}{N}\sum_{i=1}^N\mathcal{M}_i \\
	\mathcal{L}_{\mathrm{fl}}&=\frac{1}{N}\sum_{i=1}^N\frac{\mathcal{M}_i}{\mathcal{M}_{i,\mathrm{ini}}},
\end{align}	
where $\mathcal{M}_i$ represents the mismatch of an individual training waveform,
$\mathcal{M}_{i,\mathrm{ini}}$ represents the initial mismatch of the individual
training waveform, and $N$ is the total number of individual training waveforms.
Note that we choose to use two different averages, since they have different 
preferences in optimization base on waveform mismatches. For the first choice, 
simple average serves as the simplest choice of loss function, but is prone to 
be dominated by a single waveform with a large mismatch. Other waveforms with 
smaller mismatches would be insignificant comparatively, and might not be able 
to improve under such optimization. Alternatively, the second choice, normalized 
average eliminates the aforementioned issue. Nevertheless, it excludes the
information on initial mismatches. $\mathcal{L}_{\mathrm{fl}}$ restricts every
training waveform to decrease at similar rates, hence it is hard to obtain
optimized waveforms with mismatches in the same order of magnitude. Instead,
their ratios in mismatches would remain approximately the same. Conversely,
$\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust
and individual mismatches would be in a similar order of magnitude after 
optimization. In this paper, we showcase the results of using both loss functions 
and examine the differences between them.

@@

We point out that our treatment to NR waveforms is different from that of \citep{husa2016frequency, khan2016frequency}. In the original calibration process, training waveforms are hybrid waveforms of NR and SpinAlignedEOB
(SEOB) waveforms. The low frequency inspiral part is taken from the SEOB
waveforms while the rest of the waveforms are taken from NR simulations.
Instead, we solely use NR waveforms for comparison, since we are only exploring the possibility of optimizing waveform models. Thus, for simplicity sake, we ignore this procedure. On the other hand, most NR waveforms used (for both training and testing) have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, in
which they are long enough to contain part of the inspiral segment and all
merger and ringdown frequency information. We take the frequency limits as
$f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$,	
where $f_{\mathrm{RD}}$ is the frequency at ringdown. This range covers most of
the IMRPhenomD's frequency range, except the minimum frequency is set higher
than that in the original calibration due to NR length. When compared with
IMRPhenomC, the frequency range is slightly extended to have a higher maximum
frequency. We have the dimensionless frequency spacing $M\Delta f=2.5\times10^{-6}$, 
which is sufficient to capture all features of GW strain.

@@

\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
		SXS:BBH:0151 & 1.0 & -0.60    & -0.60    \\
		SXS:BBH:0001 & 1.0 &  0.00    &  0.00    \\
		SXS:BBH:0152 & 1.0 &  0.60    &  0.60    \\
		SXS:BBH:0172 & 1.0 &  0.98    &  0.98    \\
		SXS:BBH:1418 & 4.0 & -0.40    & -0.50    \\
		SXS:BBH:0167 & 4.0 &  0.00    &  0.00    \\
		SXS:BBH:1417 & 4.0 &  0.40    &  0.50    \\
		SXS:BBH:0064 & 8.0 & -0.50    & -0.46    \\
		SXS:BBH:0063 & 8.0 &  0.00    &  0.00    \\
		SXS:BBH:0065 & 8.0 &  0.50    &  0.46    \\ \midrule\bottomrule
	\end{tabularx}
	\caption{List of waveforms used to recalibrate the model. The mass ratio
	$q=m_1/m_2\geq 1$ with spins $\chi_{1,2}$. Out of the 11 waveforms listed
	here, 9 of them are also used in the original IMRPhenomD calibration.
	\citep{khan2016frequency} The two remaining waveforms were from BAM
	simulation, to which we do not have access.}
	\label{tab:q148}
\end{table}
\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0234 & 2.0 & -0.85    & -0.85    \\
		SXS:BBH:0235 & 2.0 & -0.60    & -0.60    \\
		SXS:BBH:0169 & 2.0 & 0.00     & 0.00     \\
		SXS:BBH:0256 & 2.0 & 0.60     & 0.60     \\
		SXS:BBH:0257 & 2.0 & 0.85     & 0.85     \\ \midrule\bottomrule
	\end{tabularx}
	\caption{Additional waveforms used in further recalibration.}
	\label{tab:q1248}
\end{table}

@@

With the purpose of improving downstream tasks such as parameter estimation in
mind, the more relevant metric of improvement is the distribution of improvement
in mismatch over the entire training dataset. Fig. \ref{fig:q148} shows the
distribution of log-mismatch for the testing waveforms before and after the
optimization procedure. For generality, we use a constant PSD in our loss
function. One can see the distribution have been skewed toward lower mismatch,
where the peak of the distribution is shifted by approximately an order of
magnitude, and a median mismatch is reduced by 50\%. When using
$\mathcal{L}_{\mathrm{fl}}$, we observe a similar improvement with a 22.9\%
decrease in the median.

\begin{figure}[t]
	\script{q148.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/q148.pdf}
	\caption{Distributions of waveform mismatches calculated using 
	$\mathcal{L}_{\mathrm{mean}}$. We use training waveforms listed in 
	Tab.~\ref{tab:q148} and	mismatches are weighted with the constant PSD. 
	Dashed lines represent the median of the distributions, which has decreased by 45.3\%.}
	\label{fig:q148}
\end{figure}
\begin{figure}[t]
	\script{q148_q1248_compare.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/q148_q1248_compare.pdf}
	\caption{Distributions of $\log_{10}$ difference in mismatch. The
	distribution labeled \texttt{q148} uses training waveforms listed in
	Tab.~\ref{tab:q148} while the \texttt{q1248} distribution uses waveforms
	listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}. Mismatches are calculated
	using the constant PSD with the loss function $\mathcal{L}_{\mathrm{mean}}$. 
	Dashed lines represent the median of the distributions, which has decreased by 10.4\%.}
	\label{fig:q148_q1248_compare}
\end{figure}

@@

\begin{figure}[t]
	\script{ps_q148_qchi.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/ps_q148_qchi.pdf}
	\caption{Parameter space of testing waveforms with $q$ against $\chi_{\mathrm{PN}}$. 
	We show the result from optimizing $\mathcal{L}_{\mathrm{mean}}$ with constant PSD 
	and training waveforms in Tab.~\ref{tab:q148}. Here, the colorbar represents the 
	$\log_{10}$ difference between optimized and original mismatches.}
	\label{fig:ps_q148_qchi}
\end{figure}
\begin{figure}[t]
	\script{ps_q148_chi1chi2.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/ps_q148_chi1chi2.pdf}
	\caption{Parameter space of testing waveforms with $\chi_1$ against $\chi_2$. We
	show the result from optimizing $\mathcal{L}_{\mathrm{mean}}$ with the
	constant PSD and training waveforms listed in Tab.~\ref{tab:q148}.}
	\label{fig:ps_q148}
\end{figure}

@@

We divided the parameter space into four regions to analyze the effect of the
recalibration procedure on each region separately
(Fig.~\ref{fig:ps_q148_quadrant}). The training waveforms used for fitting in
this scenario are listed in Tab.~\ref{tab:quadrants}. The top-left and
bottom-right regions have limited data for $q>4$, hence the results are only
valid up to $q\lesssim4$. From Fig.~\ref{fig:ps_q148_quadrant}, we observe that
equal-spin waveforms lying on the diagonal has great improvement. The
improvement above the diagonal is also substantial compared to recalibrate using
all waveforms at once, barring a few caused by some testing waveforms having
$q>4$. This can be seen clearer in Fig.~\ref{fig:all_quadrants}, which we see
that the optimized histogram shifts downward uniformly. On the other hand, the
improvement below the diagonal is negligible, indicating our method struggles to
further improve on top of the original IMRPhenomD waveform. Interestingly, we
can see in the top-right and bottom-left quadrant, the lower right corner shows
systematic worsening similar to the case of all waveforms. This hinted in
general, the IMRPhenomD waveform ansatz does not fit well with waveforms with
anti-aligned spin where the primary spin is positive.

\begin{figure}[t]
	\script{all_quadrants.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/all_quadrants.pdf}
	\caption{Distributions of mismatches in the top-left (Top) and bottom-right
	(Bottom) regions. We use a constant noise spectrum to calculate mismatch and
	$\mathcal{L}_{\mathrm{mean}}$ as the loss function. Waveforms in the
	top-left region generally improves while waveforms in the bottom-right
	region has very little improvement, as indicated by the median (dashed
	lines). \kw{Is the original referring to original IMRPhenomD or optimized
	using everything? Can we have a reference for optimizing using everything?}}
	\label{fig:all_quadrants}
\end{figure}

@@

In this paper, we have presented a systematic method to recalibrate GW models.
This method utilizes {\jax}'s automatic differentiation to apply
derivative-based optimization to recalibrate GW models jointly. Using the new
implementation of the IMRPhenomD model, {\ripple}, which is written in \jax, in
conjunction with NR waveforms from the SXS catalog, we recalibrate waveform
coefficients of the IMRPhenomD model. In general, the waveform accuracy can be
improved by 50\%. Comparing {\zdethp} weighted and unweighted mismatch, weighted
mismatches have a slightly better improvement. In contrast, different types of
loss function result in significantly different final mismatch distributions. As 
seen in Fig.~\ref{fig:q148}, $\mathcal{L}_{\mathrm{mean}}$ outperforms 
$\mathcal{L}_{\mathrm{fl}}$. By increasing the number of training waveforms, we see a 
slight improvement increase in Fig.~\ref{fig:q148_q1248_compare}.

>>>>

Many data analysis tasks in today's gravitational wave (GW) astrophysics, such
as match filtering search \citep{owen1996search, owen1999matched} and parameter 
estimation\citep{Dax:2021tsq, Islam:2022afg, zackay2018relative}, rely on accurate waveform
models. Because using numerical relativity (NR) waveforms in these tasks is
prohibitively expensive, the community has constructed waveform approximants
that can be evaluated much faster. There are three families of commonly used GW
approximants, which are effective-one-body (EOB) \citep{ossokine2020multipolar,
cotesta2020frequency, taracchini2014effective}, Numerical Relativity (NR)
surrogate \citep{islam2022surrogate, varma2019surrogate, varma2019surrogate2}
and phenomenological (Phenom) models \citep{husa2016frequency,
khan2016frequency, garcia2020multimode, pratten2021computationally}. While the
detail of construction of these models are different, they all have a set of
internal parameters that can be calibrated to NR waveforms. The quality of a
waveform model is determined by the ansatz used and the accuracy of the
calibrated parameters.

@@

The LIGO-VIRGO-KAGRA (LVK) collaboration
(\citep{LIGOScientific:2014pky,LIGOScientific:2021usb,LIGOScientific:2021djp,
VIRGO:2014yos,KAGRA:2020tym}.) has recently started their fourth observational
run on May 26, 2023. Because of the sensitivity improvement, the new run is
expected to double the size of current binary black hole (BBH) observations
\citep{abbott2020prospects}. The improved sensitivity also implies we expect to
detect individual events with a higher signal-to-noise ratio (SNR). This means
we can resolve more features in the signal, at the same time put a more
stringent requirement on the accuracy of our waveform model \citep{purrer2020gravitational, hu2022assessing}.

@@

Because of the large number of calibration parameters (often in a couple hundreds if not
more), waveform models is usually calibrated in pieces \citep{khan2016frequency, 
santamaria2010matching, pratten2021computationally}. This
ignores the correlation between different parts of the waveform model and limits
it quality. Recently, there has been efforts in rebuilding waveform models
\citep{khan2016frequency} that supports automatic differentiation (AD) 
\citep{ripple, Iacovelli:2022bbs, Iacovelli:2022mbg, Coogan:2022qxs}, which is a technique 
used to compute derivatives of functions down to machine precision
without issues of scaling up to high dimension or expression swelling. In
particular, {\ripple} \citep{ripple} exposes the calibration parameters to the
user. This allows us to make use of infrastructures that are heavily used in
machine learning, such as gradient descent and back propagation \cite{jax2018github, 
pytorch, tensorflow2015}, to improve the calibration of the waveform models.

% Put more focus on the joint optimization instead of having a set of more accurate 
% coefficients, since the method is the main point
In this paper, we investigate the possibility of further improving the accuracy
of a waveform model, IMRPhenomD \citep{khan2016frequency, husa2016frequency}, by jointly optimizing all the
calibration coefficients given a set of NR waveforms. Training on a few NR
waveforms, we demonstrate one can improve the match between IMRPhenomD and NR
waveforms over a decently sized parameter space, up to mass ratio $q=8$ We also
explore how different parts of the source parameter space (e.g. the primary and
second spins) respond to the optimization procedure, by optimizing the waveform
separately for different regions of the parameter space. This gives us hints on
whether the waveform model performs equally well in different regions of the
parameter space.

@@

In practice the $\Lambda$ parameters are fit for each section independently i.e., intermediate coefficients are fit whilst ignoring the MR region.
Finally, to map the grid of tuned $\Lambda$ parameters back to the intrinsic parameter space, IMRPhenomD uses the polynomial function:
\begin{align} \label{eq:Lambda}
	\Lambda^i&=\lambda_{00}^i+\lambda_{10}^i\eta \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)(\lambda_{01}^i+\lambda_{11}^i\eta+\lambda_{21}^i\eta^2) \nonumber \\ 
	&+(\chi_{\mathrm{PN}}-1)^2(\lambda_{02}^i+\lambda_{12}^i\eta+\lambda_{22}^i\eta^2) \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)^3(\lambda_{03}^i+\lambda_{13}^i\eta+\lambda_{23}^i\eta^2)\,,
\end{align}
where the $\lambda$'s are the fitting coefficients we are going to optimize below, $\eta$ is
the symmetric mass ratio, and $\chi_{\mathrm{PN}}$ is the post-Newtonian spin
parameter, which is defined as 
\begin{align}
	\chi_{\mathrm{PN}}=\frac{m_1\chi_1+m_2\chi_2}{m_1+m_2}-\frac{38\eta}{113}(\chi_1+\chi_2)\,.
\end{align}
Here, $m_{1,2}$ and $\chi_{1,2}$ are the primary and secondary mass and spin,
respectively.

@@

%\kw{Need to work on this paragraph}
Although initially independent, the stitching procedure means that each section
of the waveform intrinsically depends on the full set of $\lambda$'s. 
A slightly inaccurate set of $\lambda$'s can therefore lead to inaccuracies in
the generated waveforms. 
Thus, the calibration of these coefficients is crucial to the accuracy
of IMRPhenom GW models. 
Importantly, since the fitting was performed on the individual segments,
the final waveform is not guaranteed to have $\lambda$'s close to global minima.

@@

At the time of construction this piece-wise approach was necessary since
$\lambda$ has 209 components, making the fitting to NR simulations computationally prohibitive.
Here, for the first time we recalibrate the $\lambda$ coefficients jointly. 
This is made possible by the use of gradient-based optimization algorithms,
enabled by AD from \jax\, and {\ripple}, which are significantly more efficient in high dimensions.
% , gradients of IMRPhenomD can be easily obtained, thus
% allowing the use of gradient-based algorithms for us to recalibrate the model.

@@

Since we wish to optimize the model over the whole parameter space, we need to compare multiple model generated waveforms with NR waveforms. However, mismatch only quantifies the difference between IMRPhenomD and NR waveform for one particular set of intrinsic parameters. To take into account of various different waveforms in the parameter space, we pick waveforms from the different parts of the parameter space. We define the loss function as an average of training waveforms in two ways, the simple average of mismatches and the normalize average of mismatches,  
\begin{align}
	\mathcal{L}_{\mathrm{ave}}&=\frac{1}{N}\sum_{i=1}^N\mathcal{M}_i \\
	\mathcal{L}_{\mathrm{nl}}&=\frac{1}{N}\sum_{i=1}^N\frac{\mathcal{M}_i}{\mathcal{M}_{i,\mathrm{ini}}},
\end{align}	
where $\mathcal{M}_i$ represents the mismatch of an individual training waveform,
$\mathcal{M}_{i,\mathrm{ini}}$ represents the initial mismatch of the individual
training waveform, and $N$ is the total number of individual training waveforms.
Note that we choose to use two different averages, since they have different 
preferences in optimization base on waveform mismatches. For the first choice, 
simple average serves as the simplest choice of loss function, but is prone to 
be dominated by a single waveform with a large mismatch. Other waveforms with 
smaller mismatches would be insignificant comparatively, and might not be able 
to improve under such optimization. Alternatively, the second choice, normalized 
average eliminates the aforementioned issue. Nevertheless, it excludes the
information on initial mismatches. $\mathcal{L}_{\mathrm{nl}}$ restricts every
training waveform to decrease at similar rates, hence it is hard to obtain
optimized waveforms with mismatches in the same order of magnitude. Instead,
their ratios in mismatches would remain approximately the same. Conversely,
$\mathcal{L}_{\mathrm{ave}}$ allows the loss function to automatically adjust
and individual mismatches would be in a similar order of magnitude after 
optimization. In this paper, we showcase the results of using both loss functions 
and examine the differences between them.

@@

We point out that our treatment to NR waveforms is different from that of \citep{husa2016frequency, khan2016frequency}. In the original calibration process, training waveforms are hybrid waveforms of NR and SpinAlignedEOB
(SEOB) waveforms. The low frequency inspiral part is taken from the SEOB
waveforms while the rest of the waveforms are taken from NR simulations.
Instead, we solely use NR waveforms for comparison, since we are only exploring the possibility of optimizing waveform models. Thus, for simplicity sake, we ignore this procedure. On the other hand, most NR waveforms used (for both training and testing) have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, in
which they are long enough to contain part of the inspiral segment and all
merger and ringdown frequency information. We take the frequency limits as
$f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$,	
where $f_{\mathrm{RD}}$ is the frequency at ringdown. This range covers most of
the IMRPhenomD's frequency range, except the minimum frequency is set higher
than that in the original calibration due to NR length. When compared with
IMRPhenomC, the frequency range is slightly extended to have a higher maximum
frequency \citep{santamaria2010matching}. We have the dimensionless frequency spacing $M\Delta f=2.5\times10^{-6}$, 
which is sufficient to capture all features of GW strain.

@@

\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
		SXS:BBH:0151 & 1.0 & -0.60    & -0.60    \\
		SXS:BBH:0001 & 1.0 &  0.00    &  0.00    \\
		SXS:BBH:0152 & 1.0 &  0.60    &  0.60    \\
		SXS:BBH:0172 & 1.0 &  0.98    &  0.98    \\
		SXS:BBH:1418 & 4.0 & -0.40    & -0.50    \\
		SXS:BBH:0167 & 4.0 &  0.00    &  0.00    \\
		SXS:BBH:1417 & 4.0 &  0.40    &  0.50    \\
		SXS:BBH:0064 & 8.0 & -0.50    & -0.46    \\
		SXS:BBH:0063 & 8.0 &  0.00    &  0.00    \\
		SXS:BBH:0065 & 8.0 &  0.50    &  0.46    \\ \midrule\bottomrule
	\end{tabularx}
	\caption{List of waveforms used to recalibrate the model. The mass ratio
	$q=m_1/m_2\geq 1$ with spins $\chi_{1,2}$. Out of the 11 waveforms listed
	here, 9 of them are also used in the original IMRPhenomD calibration.
	\citep{khan2016frequency}}
	\label{tab:q148}
\end{table}
\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0234 & 2.0 & -0.85    & -0.85    \\
		SXS:BBH:0235 & 2.0 & -0.60    & -0.60    \\
		SXS:BBH:0169 & 2.0 & 0.00     & 0.00     \\
		SXS:BBH:0256 & 2.0 & 0.60     & 0.60     \\
		SXS:BBH:0257 & 2.0 & 0.85     & 0.85     \\ \midrule\bottomrule
	\end{tabularx}
	\caption{Additional waveforms used in further recalibration.}
	\label{tab:q1248}
\end{table}

@@

With the purpose of improving downstream tasks such as parameter estimation in
mind, the more relevant metric of improvement is the distribution of improvement
in mismatch over the entire training dataset. Fig. \ref{fig:q148} shows the
distribution of log-mismatch for the testing waveforms before and after the
optimization procedure. For generality, we use a constant PSD in our loss
function. One can see the distribution have been skewed toward lower mismatch,
where the peak of the distribution is shifted by approximately an order of
magnitude, and a median mismatch is reduced by 50\%. When using
$\mathcal{L}_{\mathrm{nl}}$, we observe a similar improvement with a 22.9\%
decrease in the median.

\begin{figure}[t]
	\script{q148.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/q148.pdf}
	\caption{Distributions of waveform mismatches calculated using 
	$\mathcal{L}_{\mathrm{ave}}$. We use training waveforms listed in 
	Tab.~\ref{tab:q148} and	mismatches are weighted with the constant PSD. 
	Dashed lines represent the median of the distributions, which has decreased by 45.3\%.}
	\label{fig:q148}
\end{figure}
\begin{figure}[t]
	\script{q148_q1248_compare.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/q148_q1248_compare.pdf}
	\caption{Distributions of $\log_{10}$ difference in mismatch. The
	distribution labeled \texttt{q148} uses training waveforms listed in
	Tab.~\ref{tab:q148} while the \texttt{q1248} distribution uses waveforms
	listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}. Mismatches are calculated
	using the constant PSD with the loss function $\mathcal{L}_{\mathrm{ave}}$. 
	Dashed lines represent the median of the distributions, which has decreased by 10.4\%.}
	\label{fig:q148_q1248_compare}
\end{figure}

@@

\begin{figure}[t]
	\script{ps_q148_qchi.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/ps_q148_qchi.pdf}
	\caption{Parameter space of testing waveforms with $q$ against $\chi_{\mathrm{PN}}$. 
	We show the result from optimizing $\mathcal{L}_{\mathrm{ave}}$ with constant PSD 
	and training waveforms in Tab.~\ref{tab:q148}. Here, the colorbar represents the 
	$\log_{10}$ difference between optimized and original mismatches.}
	\label{fig:ps_q148_qchi}
\end{figure}
\begin{figure}[t]
	\script{ps_q148_chi1chi2.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/ps_q148_chi1chi2.pdf}
	\caption{Parameter space of testing waveforms with $\chi_1$ against $\chi_2$. We
	show the result from optimizing $\mathcal{L}_{\mathrm{ave}}$ with the
	constant PSD and training waveforms listed in Tab.~\ref{tab:q148}.}
	\label{fig:ps_q148}
\end{figure}

@@

We divided the parameter space into four regions to analyze the effect of the
recalibration procedure on each region separately
(Fig.~\ref{fig:ps_q148_quadrant}). The training waveforms used for fitting in
this scenario are listed in Tab.~\ref{tab:quadrants} and the loss functions 
are calculated using simple average of the mismatches. The top-left and
bottom-right regions have limited data for $q>4$, hence the results are only
valid up to $q\lesssim4$. From Fig.~\ref{fig:ps_q148_quadrant}, we observe that
equal-spin waveforms lying on the diagonal has great improvement. The
improvement above the diagonal is also substantial compared to recalibrate using
all waveforms at once, barring a few caused by some testing waveforms having
$q>4$. This can be seen clearer in Fig.~\ref{fig:all_quadrants}, which we see
that the optimized histogram shifts downward uniformly. On the other hand, the
improvement below the diagonal is negligible, indicating our method struggles to
further improve on top of the original IMRPhenomD waveform. Interestingly, we
can see in the top-right and bottom-left quadrant, the lower right corner shows
systematic worsening similar to the case of all waveforms. This hinted in
general, the IMRPhenomD waveform ansatz does not fit well with waveforms with
anti-aligned spin where the primary spin is positive.

\begin{figure}[t]
	\script{all_quadrants.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/all_quadrants.pdf}
	\caption{Distributions of mismatches in the top-left (Top) and bottom-right
	(Bottom) regions. We use a constant noise spectrum to calculate mismatch and
	$\mathcal{L}_{\mathrm{ave}}$ as the loss function. Waveforms in the
	top-left region generally improves while waveforms in the bottom-right
	region has very little improvement, as indicated by the median (dashed
	lines). \kw{Is the original referring to original IMRPhenomD or optimized
	using everything? Can we have a reference for optimizing using everything?}}
	\label{fig:all_quadrants}
\end{figure}

@@

In this paper, we have presented a systematic method to recalibrate GW models.
This method utilizes {\jax}'s automatic differentiation to apply
derivative-based optimization to recalibrate GW models jointly. Using the new
implementation of the IMRPhenomD model, {\ripple}, which is written in \jax, in
conjunction with NR waveforms from the SXS catalog, we recalibrate waveform
coefficients of the IMRPhenomD model. In general, the waveform accuracy can be
improved by 50\%. Comparing {\zdethp} weighted and unweighted mismatch, weighted
mismatches have a slightly better improvement. In contrast, different types of
loss function result in significantly different final mismatch distributions. As 
seen in Fig.~\ref{fig:q148}, $\mathcal{L}_{\mathrm{ave}}$ outperforms 
$\mathcal{L}_{\mathrm{nl}}$. By increasing the number of training waveforms, we see a 
slight improvement increase in Fig.~\ref{fig:q148_q1248_compare}.
